Tulai

### 2023-09-05 Tulai.qmd
aa

---
title: "Tulai"
author: "Setareh Shafizadeh, Ben Marwick"
date: now
format: 
  html:
    self-contained: true
execute: 
  warning: false
  fig-dpi: 600
title-block-published: "last updated"
editor: visual
---

## Introduction

@@

  # Separate 'TA' column into 'area' and 'depth' columns
  separate(TA, into = c("area", 
                        "depth"), 
           sep = "/",
           remove = FALSE,
           convert = TRUE) %>%

  # Convert 'depth' to numeric
  mutate(depth = as.numeric(depth)) %>%

@@

# create 'layer' column
  mutate(layer = case_when(
      area == "tp1" & depth == "0-10 cm" ~ "c",
      area == "tp1" & depth == "0-20 cm" ~ "c",
      area == "tp1" & depth == "10-30 cm" ~ "c",
      area == "tp1" & depth == "10-90 cm" ~ "c",
    area == "tp1" & depth == "20-30 cm" ~ "c",
    area == "tp1" & depth == "30-40 cm" ~ "c",
    area == "tp1" & depth == "40-50 cm" ~ "c",
    area == "tp1" & depth == "50-100 cm" ~ "c",
    area == "tp1" & depth == "50-60 cm" ~ "c",
    area == "tp1" & depth == "50-70 cm" ~ "c",
    area == "tp1" & depth == "80-90 cm" ~ "c",
    area == "tp1" & depth == "90-100 cm" ~ "c",
    area == "tp1" & depth == "100-110 cm" ~ "c",
    area == "tp1" & depth == "110-120 cm" ~ "c",
    area == "tp1" & depth == "120-130 cm" ~ "c",
    area == "tp1" & depth == "120-140 cm" ~ "c",
    area == "d1" & depth == "0-20 cm" ~ "c",
    area == "d1" & depth == "20-30 cm" ~ "c",
    area == "d1" & depth == "40-60 cm" ~ "a",
    area == "d1" & depth == "50-60 cm" ~ "a",
    area == "d1" & depth == "60-70 cm" ~ "a",
    area == "d1" & depth == "70-80 cm" ~ "a",
    area == "ta" & depth == "0-30 cm" ~ "e",
    area == "ta" & depth == "30-40 cm" ~ "e",
    area == "ta" & depth == "40-50 cm" ~ "e or d",
    area == "ta" & depth == "50-60 cm" ~ "e or d",
    area == "ta" & depth == "60-70 cm" ~ "e or d",
    area == "ta" & depth == "70-120 cm" ~ "e or d",
    area == "ta" & depth == "120-150 cm" ~ "d",
    area == "ta" & depth == "150-180 cm" ~ "d",
    area == "ta" & depth == "180-190 cm" ~ "d",
    area == "ta" & depth == "190-200 cm" ~ "d",
    area == "ta" & depth == "200-210 cm" ~ "d",
    area == "ta" & depth == "210-220 cm" ~ "d",
    area == "ta" & depth == "220-230 cm" ~ "d",
    area == "ta" & depth == "230-240 cm" ~ "d",
    area == "ta" & depth == "230-250 cm" ~ "d",
    area == "ta" & depth == "250-260 cm" ~ "d",
    area == "ta" & depth == "260-270 cm" ~ "d",

@@

tl2 <- tl2 %>%
  mutate(
    Blank = str_to_lower(Blank),
    Blank = str_trim(Blank),
    Blank = str_replace_all(Blank, "[./]", "-"),
    Blank = str_replace_all(Blank, "\\s*-\\s*", "-")
  ) %>%
  mutate(Blank = case_when(
             Blank == "peo" ~ "bladelet-pro",
             Blank == "pro" ~ "bladelet-pro",
             Blank == "bladelet-dis" ~ "bladelet-dis",
             Blank == "microblade" ~ "bladelet-complete",
             Blank == "microblade-pro" ~ "bladelet-pro",
             Blank == "microblade-med" ~ "bladelet-med",
             Blank == "micrpblade-dis" ~ "bladelet-dis",
             Blank == "dis" ~ "bladelet-dis",
             Blank == "bladelrt-med" ~ "bladelet-med",
             Blank == "bladelet-mes" ~ "bladelet-med",
             Blank == "bladelet-,ed" ~ "bladelet-med",
             Blank == "bladelet-bladelet-med" ~ "bladelet-med",
             Blank == "medial" ~ "bladelet-med",
             Blank == "bladelet-bladelet-pro" ~ "bladelet-pro",
             Blank == "bladelert-pro" ~ "bladelet-pro",
             Blank == "microblade-medial" ~ "bladelet-med",
             Blank == "nicroblade-pro" ~ "bladelet-pro",
             Blank == "microblade-?" ~ "bladelet-complete",
             Blank == "med" ~ "bladelet-med",
             Blank == "bladeler-pro" ~ "bladelet-pro",
             Blank == "microblade-dis" ~ "bladelet-dis",
             Blank == "microblde-dis" ~ "bladelet-dis",
             Blank == "bladlet-pro" ~ "bladelet-pro",
             Blank == "indistinct" ~ "bladelet-med",
             Blank == "flake" ~ "flake-complete",
             Blank == "dladelet-pro" ~ "bladelet-pro",
             Blank == "dladelet-pro" ~ "bladelet-pro",
             Blank == "bldelet-bladelet-pro" ~ "bladelet-pro",
             Blank == "bldelet-pro" ~ "bladelet-pro",
             Blank == "nicroblade-pro" ~ "bladelet-pro",
             Blank == "microblade-?" ~ "bladelet-med",
             Blank == "bladelt-pro" ~ "bladelet-pro",
             Blank == "bladeket-med" ~ "bladelet-med",
             Blank == "nicroblade-pro" ~ "bladelet-pro",
             Blank == "blaelet" ~ "bladelet-complete",
             Blank == "bladele-med" ~ "bladelet-med",
             Blank == "bladelet-nearly complete" ~ "bladelet-complete",
             Blank == "blade-nearly complete" ~ "blade-complete",
             Blank == "bladelet-?" ~ "bladelet-complete",
             Blank == "flake?" ~ "flake-complete",
             Blank == "micrpblade-dis" ~ "bladelet-dis",
             Blank == "blaelet-med" ~ "bladelet-med",
             Blank == "blaedlet-pro" ~ "bladelet-pro",  
             Blank == "bladelt-pro" ~ "bladelet-pro",   
             Blank == "bladelet-nearly complete" ~ "bladelet-complete", 
             Blank == "microblade-med" ~ "bladelet-med", 
             Blank == "bldelet-bladelet-pro" ~ "bladelet-pro",
             Blank == "bldelet-flake" ~ "blade-complete",
             Blank == "mwdial" ~ "bladelet-med",
             Blank == "blade" ~ "blade-complete",
             Blank == "bladelt" ~ "bladelet-complete",
             Blank == "bladlet" ~ "bladelet-complete",
             Blank == "bladlet-med" ~ "bladelet-med",
             Blank == "bladelt-med" ~ "bladelet-med",
             Blank == "flaje" ~ "flake",
             Blank == "flke" ~ "flake",
             Blank == "bladelet" ~ "bladelet-complete",
             Blank == "NA" ~ "bladelet-",
             Blank == "thick flake" ~ "flake",
             Blank == "flke-mid" ~ "flake-med",
             Blank == "flake-blade" ~ "flake",
             Blank == "angular flake?" ~ "flake",
             Blank == "flake-prox" ~ "flake-pro",
             Blank == "fklae-dis" ~ "flake-dis",
             Blank == "NA" ~ NA_character_,
    TRUE ~ Blank
  )) %>%
  mutate(Blank = na_if(Blank, ""))

@@

```{r}
# Remove rows 
tl3 <- tl3[!(rownames(tl3) %in% c("1030", "644", "543")), ]
```

@@

# Check
print(unique(tl4$SHINE))

@@

# Check 
unique(tl4$`core-typology`)
unique(tl4$`core-technology`)
```

@@

```{r}
# First convert all columns to character type (cant remember why I did that, )
tl4 <- tl4 %>%
  mutate(across(everything(), as.character))

@@

@@

 select(lithic_id, 
         TA, 
         area, 
         layer,
        cortex,
         depth, upper, lower, midpoint, 
         `length`, `width`, `thickness`, 
         typology, 
        utilization, 
        `sickle shine`,
        retouch, 
        `typology-tool-core`,
        `tool-typology`,
         blank, `blank part`,

@@

```{r}
# Update the 'typology-tool-core' column based on the condition in 'Retouch'
tl_final <- tl_final %>% 
  mutate(
    `typology-tool-core` = case_when(
      retouch == 1 ~ "tool",           
      utilization == 1 ~ "tool",        
      TRUE ~ `typology-tool-core`      
    )
  )

@@

```{r}
tl_final <- tl_final %>% 
  mutate(
    `tool-typology` = case_when(
      utilization == 1 ~ "utilized tool",  
      TRUE ~ `tool-typology`               
    )
  )

@@

     # If Retouch is 1 and tool-typology is NA
       retouch == 1 & is.na(`tool-typology`) ~ "retouched piece",

@@

@@

@@

tl_final <- tl_final %>%

  mutate(`core-fragment` = if_else(`core-fragment` == "", NA_character_, `core-fragment`)) %>%
  # Convert "pyramid-percussion" to "pyramid" in `core-fragment` column
  mutate(`core-fragment` = 
           if_else(
             `core-fragment` == "pyramid-percussion", "pyramid", `core-fragment`)) %>%
  # Update existing `core` column with selected values from `core-fragment` column
  mutate(core = if_else(`core-fragment` %in% 
                          c("pyramid", 
                            "cylinder/prismatic", 
                            "flat", 
                            "bullet", 
                            "core on flake", 
                            "pyramid"), 
                        `core-fragment`, 
                        core))

@@

@@

tl_final <- tl_final %>%
  # Transfer 'rejuvenation piece' and 'core fragment' from `core-typology` to `corefrag`
  mutate(corefrag = case_when(
    `core-typology` == "rejuvenation piece" ~ "rejuvenation piece",
    `core-typology` == "core fragment" ~ "core fragment",
    TRUE ~ corefrag
  )) %>%
  # Remove 'rejuvenation piece' and 'core fragment' from `core-typology` 
  mutate(`core-typology` = if_else(`core-typology` %in% c("rejuvenation piece", "core fragment"), NA_character_, `core-typology`))

unique(tl_final$`core-typology`)
unique(tl_final$`corefrag`)

@@

# Update the 'blank' column based on conditions in the 'core-fragment' column

tl_final <- tl_final %>%
  mutate(blank = if_else(`core-fragment` %in% c(
    "core fragment", "core tablet", "rejuvenation piece", "pyramid", "crested bladelet", 
    "shapeless", "primary flake", "flat", "bullet", "core on flake", "cylinder/prismatic",
    "primary blade", "rejuvenation piece-side", "core platform rejuvention"),
    NA_character_, 
    blank))

>>>>

---
title: "Tulai"
author: "S.SH"
date: now
format: 
  html:
    self-contained: true
execute: 
  warning: false
  fig-dpi: 600
title-block-published: "last updated"
editor: visual
---

@@

  # Separate 'TA' column into 'area' and 'depth' columns
  separate(TA, into = c("area", 
                        "depth"), 
           sep = "/",
           remove = FALSE,
           convert = TRUE) %>%
  # Convert 'depth' to numeric
  mutate(depth = as.numeric(depth)) %>%

@@

# create 'layer' column
  mutate(layer = case_when(
      area == "tp1" & depth == "0-10 cm" ~ "c",
      area == "tp1" & depth == "0-20 cm" ~ "c",
      area == "tp1" & depth == "10-30 cm" ~ "c",
      area == "tp1" & depth == "10-90 cm" ~ "c",
    area == "tp1" & depth == "20-30 cm" ~ "c",
    area == "tp1" & depth == "30-40 cm" ~ "c",
    area == "tp1" & depth == "40-50 cm" ~ "c",
    area == "tp1" & depth == "50-100 cm" ~ "c",
    area == "tp1" & depth == "50-60 cm" ~ "c",
    area == "tp1" & depth == "50-70 cm" ~ "c",
    area == "tp1" & depth == "80-90 cm" ~ "c",
    area == "tp1" & depth == "90-100 cm" ~ "c",
    area == "tp1" & depth == "100-110 cm" ~ "c",
    area == "tp1" & depth == "110-120 cm" ~ "c",
    area == "tp1" & depth == "120-130 cm" ~ "c",
    area == "tp1" & depth == "120-140 cm" ~ "c",
    area == "d1" & depth == "0-20 cm" ~ "c",
    area == "d1" & depth == "20-30 cm" ~ "c",
    area == "d1" & depth == "40-60 cm" ~ "a",
    area == "d1" & depth == "50-60 cm" ~ "a",
    area == "d1" & depth == "60-70 cm" ~ "a",
    area == "d1" & depth == "70-80 cm" ~ "a",
    area == "ta" & depth == "0-30 cm" ~ "e",
    area == "ta" & depth == "30-40 cm" ~ "e",
    area == "ta" & depth == "40-50 cm" ~ "e",
    area == "ta" & depth == "50-60 cm" ~ "e",
    area == "ta" & depth == "60-70 cm" ~ "e",
    area == "ta" & depth == "70-120 cm" ~ "d",
    area == "ta" & depth == "120-150 cm" ~ "d",
    area == "ta" & depth == "150-180 cm" ~ "d",
    area == "ta" & depth == "180-190 cm" ~ "d",
    area == "ta" & depth == "190-200 cm" ~ "d",
    area == "ta" & depth == "200-210 cm" ~ "d",
    area == "ta" & depth == "210-220 cm" ~ "d",
    area == "ta" & depth == "220-230 cm" ~ "d",
    area == "ta" & depth == "230-240 cm" ~ "d",
    area == "ta" & depth == "230-250 cm" ~ "d",
    area == "ta" & depth == "250-260 cm" ~ "d",
    area == "ta" & depth == "260-270 cm" ~ "d",

@@

tl2 <- tl2 %>%
  mutate(
    Blank = str_to_lower(Blank),
    Blank = str_trim(Blank),
    Blank = str_replace_all(Blank, "[./]", "-"),
    Blank = str_replace_all(Blank, "\\s*-\\s*", "-")
  ) %>%
  mutate(Blank = case_when(
             Blank == "peo" ~ "bladelet-pro",
             Blank == "pro" ~ "bladelet-pro",
             Blank == "bladelet-dis" ~ "bladelet-dis",
             Blank == "microblade" ~ "bladelet-complete",
             Blank == "microblade-pro" ~ "bladelet-pro",
             Blank == "microblade-med" ~ "bladelet-med",
             Blank == "micrpblade-dis" ~ "bladelet-dis",
             Blank == "dis" ~ "bladelet-dis",
             Blank == "bladelrt-med" ~ "bladelet-med",
             Blank == "bladelet-mes" ~ "bladelet-med",
             Blank == "bladelet-,ed" ~ "bladelet-med",
             Blank == "bladelet-bladelet-med" ~ "bladelet-med",
             Blank == "medial" ~ "bladelet-med",
             Blank == "bladelet-bladelet-pro" ~ "bladelet-pro",
             Blank == "bladelert-pro" ~ "bladelet-pro",
             Blank == "microblade-medial" ~ "bladelet-med",
             Blank == "nicroblade-pro" ~ "bladelet-pro",
             Blank == "microblade-?" ~ "bladelet-complete",
             Blank == "med" ~ "bladelet-med",
             Blank == "bladeler-pro" ~ "bladelet-pro",
             Blank == "microblade-dis" ~ "bladelet-dis",
             Blank == "microblde-dis" ~ "bladelet-dis",
             Blank == "bladlet-pro" ~ "bladelet-pro",
             Blank == "indistinct" ~ "bladelet-med",
             Blank == "flake" ~ "flake-complete",
             Blank == "dladelet-pro" ~ "bladelet-pro",
             Blank == "dladelet-pro" ~ "bladelet-pro",
             Blank == "bldelet-bladelet-pro" ~ "bladelet-pro",
             Blank == "bldelet-pro" ~ "bladelet-pro",
             Blank == "nicroblade-pro" ~ "bladelet-pro",
             Blank == "microblade-?" ~ "bladelet-med",
             Blank == "bladelt-pro" ~ "bladelet-pro",
             Blank == "bladeket-med" ~ "bladelet-med",
             Blank == "nicroblade-pro" ~ "bladelet-pro",
             Blank == "blaelet" ~ "bladelet-complete",
             Blank == "bladele-med" ~ "bladelet-med",
             Blank == "bladelet-nearly complete" ~ "bladelet-complete",
             Blank == "blade-nearly complete" ~ "blade-complete",
             Blank == "bladelet-?" ~ "bladelet-complete",
             Blank == "flake?" ~ "flake-complete",
             Blank == "micrpblade-dis" ~ "bladelet-dis",
             Blank == "blaelet-med" ~ "bladelet-med",
             Blank == "blaedlet-pro" ~ "bladelet-pro",  
             Blank == "bladelt-pro" ~ "bladelet-pro",   
             Blank == "bladelet-nearly complete" ~ "bladelet-complete", 
             Blank == "microblade-med" ~ "bladelet-med", 
             Blank == "bldelet-bladelet-pro" ~ "bladelet-pro",
             Blank == "bldelet-flake" ~ "blade-complete",
             Blank == "mwdial" ~ "bladelet-med",
             Blank == "blade" ~ "blade-complete",
             Blank == "bladelt" ~ "bladelet-complete",
             Blank == "bladlet" ~ "bladelet-complete",
             Blank == "bladlet-med" ~ "bladelet-med",
             Blank == "bladelt-med" ~ "bladelet-med",
             Blank == "flaje" ~ "flake",
             Blank == "flke" ~ "flake",
             Blank == "bladelet" ~ "bladelet-complete",
             Blank == "NA" ~ "bladelet-",
             Blank == "thick flake" ~ "flake",
             Blank == "flke-mid" ~ "flake-med",
             Blank == "flake-blade" ~ "flake",
             Blank == "angular flake?" ~ "flake",
             Blank == "flake-prox" ~ "flake-pro",
             Blank == "fklae-dis" ~ "flake-dis",
             Blank == "bladelete" ~ "bladelet",
             Blank == "NA" ~ NA_character_,
    TRUE ~ Blank
  )) %>%
  mutate(Blank = na_if(Blank, ""))

@@

```{r}
# Remove rows based on row names
tl3 <- tl3[!(rownames(tl3) %in% c("1030", "644", "543")), ]
```

@@

# Check unique values in the modified columns
print(unique(tl4$SHINE))

@@

# Check the unique values again
unique(tl4$`core-typology`)
unique(tl4$`core-technology`)
```

@@

```{r}
# First convert all columns to character type
tl4 <- tl4 %>%
  mutate(across(everything(), as.character))

@@

```{r}
tl_final <- tl_final %>%
  mutate(al = paste0(toupper(area), "_", layer)) %>%
  relocate(al, .after = layer)

@@

 select(lithic_id, 
         TA, 
         area, 
         layer,
        al,
        cortex,
         depth, upper, lower, midpoint, 
         `length`, `width`, `thickness`, 
         typology, 
        utilization, 
        `sickle shine`,
        retouch, 
        `typology-tool-core`,
        `tool-typology`,
         blank, `blank part`,

@@

```{r}
# Update the 'typology-tool-core' column based on the condition in 'Retouch'
tl_final <- tl_final %>% 
  mutate(
    `typology-tool-core` = case_when(
      retouch == 1 ~ "tool",           
      utilization == 1 ~ "tool",
      `sickle shine` == 1 ~ "tool",

      TRUE ~ `typology-tool-core`      
    )
  )

@@

```{r}
```

tl_final <- tl_final %>% 
  mutate(
    `tool-typology` = case_when(
      utilization == 1 ~ "utilized tool",
      `sickle shine` == 1 ~ "sickle shine",
      TRUE ~ `tool-typology`               
    )
  )

@@

     # If Retouch is 1 and tool-typology is NA
       retouch == 1 & is.na(`tool-typology`) ~ "retouched piece",
       `sickle shine` == 1 & is.na(`tool-typology`) ~ "sickle shine",

@@

# Show updated data frame
unique(tl_final$`tool-typology`)

@@

                   #DO NOT RUN#############

tl_final <- tl_final %>%
  mutate(
    `core-fragments` = str_to_lower(`core-fragment`),
    `core-fragments` = str_trim(`core-fragment`),
    `core-fragments` = str_replace_all(`core-fragment`, "^[-._/NA\\s]*$|^\\s*$", "0")
  ) %>%
  mutate(`core-fragment` = case_when(

            `core-fragment` == "rejuvenation piece" ~ "rejuvenation piece"

            `core-fragment` == "crested bladelet" ~ "crested bladelet",

            `core-fragment` == "crested bladelet" ~ "crested bladelet",

            `core-fragment` == "primaryflake" ~ "primary flake",
            `core-fragment` == "primary flake" ~ "primary flake",

            `core-fragment` == "primary blade" ~ "primary blade",

            `core-fragment` == "rejuvenationpiece-side" ~ "rejuvenation piece-side",
            `core-fragment` == "rejuvention piece-side" ~ "rejuvenation piece-side",

            `core-fragment` == "multidirectional core fragment" ~ "shapeless",
            `core-fragment` == "multidirectionalcore fragment" ~ "shapeless",
            `core-fragment` == "multidirectionalbladelet core" ~ "shapeless",

            `core-fragment` == ("bipolar-percussion") ~ "cylinder/prismatic-percussion",

            `core-fragment` == "core platform rejuvention" ~ "core tablet",
            `core-fragment` == "bipolar" ~ "cylinder/prismatic",
             `core-fragment` == "coretablet" ~ "core tablet",
            `core-fragment` == "crestedbladelet" ~ "crested piece",
            `core-fragment` == "coreon flake" ~ "core on flake",
            `core-fragment` == "corefragment-pyramid" ~ "pyramid",
            `core-fragment` == "corepreparation piece-platform" ~ "core tablet",

    TRUE ~ `core-fragment`
  ))

unique(`tl_final`$`core-fragment`)
```

@@

tl_final <- tl_final %>%
  # Convert empty strings to NA in `core-fragment` column
  mutate(`core-fragment` = if_else(`core-fragment` == "", NA_character_, `core-fragment`)) %>%
  # Convert "pyramid-percussion" to "pyramid" in `core-fragment` column
  mutate(`core-fragment` = 
           if_else(
             `core-fragment` == "pyramid-percussion", "pyramid", `core-fragment`)) %>%
  # Update existing `core` column with selected values from `core-fragment` column
  mutate(core = if_else(`core-fragment` %in% 
                          c("pyramid", 
                            "cylinder/prismatic", 
                            "flat", 
                            "bullet", 
                            "core on flake", 
                            "pyramid"), 
                        `core-fragment`, 
                        core))

unique(tl_final$`core-fragment`)
unique(tl_final$`corefrag`)
unique(tl_final$`core`)
```

unique(tl_final$`core-fragment`)
unique(tl_final$`corefrag`)
unique(tl_final$`core`)
```

# Process the tl_final_area dataframe using pipe (%>%)
tl_final <- tl_final %>%
  # Transfer 'rejuvenation piece' and 'core fragment' from `core-typology` to `corefrag`
  mutate(corefrag = case_when(
    `core-typology` == "rejuvenation piece" ~ "rejuvenation piece",
    `core-typology` == "core fragment" ~ "core fragment",
    TRUE ~ corefrag
  )) %>%
  # Remove 'rejuvenation piece' and 'core fragment' from `core-typology` but keep the column
  mutate(`core-typology` = if_else(`core-typology` %in% c("rejuvenation piece", "core fragment"), NA_character_, `core-typology`))

unique(tl_final$`core-typology`)
unique(tl_final$`corefrag`)
unique(tl_final$`core-fragment`)

@@

# Update the 'blank' column based on conditions in the 'core-fragment' column
tl_final <- tl_final %>%
  mutate(blank = if_else(`core-fragment` %in% c(
    "core fragment", "core tablet", "rejuvenation piece", "pyramid", "crested bladelet", 
    "shapeless", "primary flake", "flat", "bullet", "core on flake", "cylinder/prismatic",
    "primary blade", "rejuvenation piece-side", "core platform rejuvention"),
    NA_character_, 
    blank))

# Show the updated data
print(tl_final)
unique(tl_final$`core-fragment`)
unique(tl_final$`blank`)

Summary tables: 
1.AREA
```{r}
# Create a summary table
summary_table_area <- tl_final %>%
  group_by(area) %>%
  summarise(
    `Cores (n)` = sum(!is.na(core) & core != ""),
    `Core rejuvenations pieces (n)` = sum(!is.na(corefrag) & corefrag != ""),
    `Blade (n)` = sum(blank == "blade", na.rm = TRUE),
    `Flake (n)` = sum(blank == "flake", na.rm = TRUE),
    `Bladelet (n)` = sum(blank == "bladelet", na.rm = TRUE),
    `Retouched Tools (n)` = sum(retouch == 1, na.rm = TRUE),
    `Utilized (n)` = sum(utilization == 1, na.rm = TRUE),
    `Sickle Shine (n)` = sum(`sickle shine` == 1, na.rm = TRUE)
  )

# Show summary table
print(summary_table_area)

```{r}
library(kableExtra)

kable(summary_table_area, 
      caption = "Table: Summary by Area", 
      align = 'c') %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                full_width = F, 
                font_size = 12) %>%
  row_spec(0, bold = TRUE, font_size = 14)

2.LAYER
```{r}
# Create a summary table
summary_table_layer <- tl_final %>%
  group_by(layer) %>%  # Note that 'area' is replaced with 'layer'
  summarise(
    `Cores (n)` = sum(!is.na(core)),  # Count non-NA core entries
    `Core rejuvenations pieces (n)` = sum(!is.na(corefrag)),  # Count non-NA corefrag entries
    `Blade (n)` = sum(blank == "blade", na.rm = TRUE),
    `Flake (n)` = sum(blank == "flake", na.rm = TRUE),
    `Bladelet (n)` = sum(blank == "bladelet", na.rm = TRUE),
    `Retouched Tools (n)` = sum(retouch == 1, na.rm = TRUE),
    `Utilized (n)` = sum(utilization == 1, na.rm = TRUE),
    `Sickle Shine (n)` = sum(`sickle shine` == 1, na.rm = TRUE)
  )

```{r}
# Generate the table
kable(summary_table_layer, 
      caption = "Table: Summary by Layer", 
      align = 'c') %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                full_width = F, 
                font_size = 12) %>%
  row_spec(0, bold = TRUE, font_size = 14)

AREA_LAYER:
```{r}
# Create a summary table based on the 'al' column
summary_table_al <- tl_final %>%
  group_by(al) %>%  
  summarise(
    `Cores (n)` = sum(!is.na(core)),  # Count non-NA core entries
    `Core rejuvenations pieces (n)` = sum(!is.na(corefrag)),  # Count non-NA corefrag entries
    `Blade (n)` = sum(blank == "blade", na.rm = TRUE),
    `Flake (n)` = sum(blank == "flake", na.rm = TRUE),
    `Bladelet (n)` = sum(blank == "bladelet", na.rm = TRUE),
    `Retouched Tools (n)` = sum(retouch == 1, na.rm = TRUE),
    `Utilized (n)` = sum(utilization == 1, na.rm = TRUE),
    `Sickle Shine (n)` = sum(`sickle shine` == 1, na.rm = TRUE)
  )

```{r}
# Generate the table
library(kableExtra)

kable(summary_table_al, 
      caption = "Table: Summary by Al", 
      align = 'c') %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                full_width = F, 
                font_size = 12) %>%
  row_spec(0, bold = TRUE, font_size = 14)

DEPTH:

```{r}
tl_final <- tl_final %>%
  mutate(depth = gsub("cm", "", depth))
```

```{r}
# Create a summary table based on the 'depth' column
summary_table_depth <- tl_final %>%
  group_by(depth) %>%  
  summarise(
    `Cores (n)` = sum(!is.na(core)),  # Count non-NA core entries
    `Core rejuvenations pieces (n)` = sum(!is.na(corefrag)),  # Count non-NA corefrag entries
    `Blade (n)` = sum(blank == "blade", na.rm = TRUE),
    `Flake (n)` = sum(blank == "flake", na.rm = TRUE),
    `Bladelet (n)` = sum(blank == "bladelet", na.rm = TRUE),
    `Retouched Tools (n)` = sum(retouch == 1, na.rm = TRUE),
    `Utilized (n)` = sum(utilization == 1, na.rm = TRUE),
    `Sickle Shine (n)` = sum(`sickle shine` == 1, na.rm = TRUE)
  )

```{r}
# Generate the table
library(kableExtra)

kable(summary_table_depth, 
      caption = "Table: Summary by Depth", 
      align = 'c') %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                full_width = F, 
                font_size = 12) %>%
  row_spec(0, bold = TRUE, font_size = 14)

```{r}
# Create a summary table based on the 'depth', 'area', and 'layer' columns
summary_table_depth_area_layer <- tl_final %>%
  group_by(depth, area, layer, upper, lower) %>%  
  summarise(
    `Cores (n)` = sum(!is.na(core)),  # Count non-NA core entries
    `Core rejuvenations pieces (n)` = sum(!is.na(corefrag)),  # Count non-NA corefrag entries
    `Blade (n)` = sum(blank == "blade", na.rm = TRUE),
    `Flake (n)` = sum(blank == "flake", na.rm = TRUE),
    `Bladelet (n)` = sum(blank == "bladelet", na.rm = TRUE),
    `Retouched Tools (n)` = sum(retouch == 1, na.rm = TRUE),
    `Utilized (n)` = sum(utilization == 1, na.rm = TRUE),
    `Sickle Shine (n)` = sum(`sickle shine` == 1, na.rm = TRUE)
  )

# Generate the table
library(kableExtra)

kable(summary_table_depth_area_layer, 
      caption = "Table: Summary by Depth, Area, and Layer", 
      align = 'c') %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                full_width = F, 
                font_size = 12) %>%
  row_spec(0, bold = TRUE, font_size = 14)

Techno-Typo:

```{r}
# Load dplyr library
library(dplyr)

# Prepare data for Summary table-area
summary_table_area <- tl_final_area %>% 
  filter(area %in% c("ta", "d1", "tp1")) %>%
  group_by(area) %>% 
  summarise(
    blade_count = sum(blank == "blade", na.rm = TRUE),
    flake_count = sum(blank == "flake", na.rm = TRUE),
    bladelet_count = sum(blank == "bladelet", na.rm = TRUE),
    tool_count = sum(retouch == 1, na.rm = TRUE) + sum(utilization == 1, na.rm = TRUE),
    core_count = sum(!is.na(`core-typology`), na.rm = TRUE),
    core_rejuvenation_count = sum(!is.na(`corefrag`), na.rm = TRUE),
    retouched_tool_count = sum(retouch == 1, na.rm = TRUE),
    utilized_tool_count = sum(utilization == 1, na.rm = TRUE),
    shine_tool_count = sum(`sickle shine` == 1, na.rm = TRUE),
    total = n()
  )

# Adding a new row with total counts
# Adding a new row with total counts
summary_table_area <- summary_table_area %>% 
  add_row(
    area = "Total",
    blade_count = sum(summary_table_area$blade_count, na.rm = TRUE),
    flake_count = sum(summary_table_area$flake_count, na.rm = TRUE),
    bladelet_count = sum(summary_table_area$bladelet_count, na.rm = TRUE),
    tool_count = sum(summary_table_area$tool_count, na.rm = TRUE),
    core_count = sum(summary_table_area$core_count, na.rm = TRUE),
    core_rejuvenation_count = sum(summary_table_area$core_rejuvenation_count, na.rm = TRUE),
    retouched_tool_count = sum(summary_table_area$retouched_tool_count, na.rm = TRUE),
    utilized_tool_count = sum(summary_table_area$utilized_tool_count, na.rm = TRUE),
    shine_tool_count = sum(summary_table_area$shine_tool_count, na.rm = TRUE),
    total = sum(summary_table_area$total, na.rm = TRUE)  # Explicitly setting the total
  )

```{r}
# Prepare data for Summary table-al
summary_table_al <- tl_final_area %>% 
  group_by(al) %>% 
  summarise(
    blade_count = sum(blank == "blade", na.rm = TRUE),
    flake_count = sum(blank == "flake", na.rm = TRUE),
    bladelet_count = sum(blank == "bladelet", na.rm = TRUE),
    tool_count = sum(retouch == 1, na.rm = TRUE) + sum(utilization == 1, na.rm = TRUE),
    core_count = sum(!is.na(`core-typology`), na.rm = TRUE),
    core_rejuvenation_count = sum(!is.na(`corefrag`), na.rm = TRUE),
    retouched_tool_count = sum(retouch == 1, na.rm = TRUE),
    utilized_tool_count = sum(utilization == 1, na.rm = TRUE),
    shine_tool_count = sum(`sickle shine` == 1, na.rm = TRUE),
    total = n()
  )

# Adding a new row with total counts
summary_table_al <- summary_table_al %>% 
  add_row(
    al = "Total",
    blade_count = sum(summary_table_al$blade_count, na.rm = TRUE),
    flake_count = sum(summary_table_al$flake_count, na.rm = TRUE),
    bladelet_count = sum(summary_table_al$bladelet_count, na.rm = TRUE),
    tool_count = sum(summary_table_al$tool_count, na.rm = TRUE),
    core_count = sum(summary_table_al$core_count, na.rm = TRUE),
    core_rejuvenation_count = sum(summary_table_al$core_rejuvenation_count, na.rm = TRUE),
    retouched_tool_count = sum(summary_table_al$retouched_tool_count, na.rm = TRUE),
    utilized_tool_count = sum(summary_table_al$utilized_tool_count, na.rm = TRUE),
    shine_tool_count = sum(summary_table_al$shine_tool_count, na.rm = TRUE),
    total = sum(summary_table_al$total, na.rm = TRUE)  # Explicitly setting the total
  )

# View the summary table
print(summary_table_al)

```{r}
summary_table_al <- summary_table_al %>%
  filter(!(al %in% c("D1_NA", "TP1_NA")))

kable(summary_table_al, 
      col.names = c("Area-Layer", "Blade", "Flake", "Bladelet", "Tools", "Cores", "Core Rejuvenations", "Retouched Tools", "Utilized Tools", "Sickle Shine", "Total"), 
      caption = "Table: Summary ", 
      align = 'c') %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                full_width = FALSE, 
                font_size = 12) %>%
  row_spec(0, bold = TRUE, font_size = 14)

library(knitr)
library(dplyr)
# Rename columns
summary_table_area <- summary_table_area %>%
  rename(
    "Area" = area,
    "Blade" = blade_count,
    "Flake" = flake_count,
    "Bladelet" = bladelet_count,
    "Retouched Tools" = retouched_tool_count,
    "Tools" = tool_count,
    "Utilized Tools" = utilized_tool_count,
    "Sickle Shine" = shine_tool_count,
    "Cores" = core_count,
    "Core rejuvenations pieces" = core_rejuvenation_count,
    "Total" = total
  ) %>%
  mutate(
    Area = if_else(row_number() == n() & is.na(Area), "Total", Area)
  ) %>%
  mutate(
    Area = recode(Area, "ta" = "TA", "d1" = "D1", "tp1" = "TP1", "Total" = "Total")
  )
library(kableExtra)

kable(summary_table_area, 
      caption = "Table 1: Count of debitage type and tools in three different main areas.", 
      align = 'c') %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), font_size = 12) %>%

```{r}
# summary table-core and core rejuvenation
```

```{r}
# summary table for tools: retouched-utelized-sickle

@@

# Create a summary table
summary_table_tool_area <- tl_final_area %>%
  filter(!is.na(tooltypegroup), tooltypegroup != "") %>%  # Exclude rows where 'tooltypegroup' is NA or empty
  group_by(area, tooltypegroup) %>%  # Group the data by area and tooltypegroup
  tally() %>%  # Count the number of rows in each group
  arrange(area, tooltypegroup) %>%  # Arrange by area and tooltypegroup
  spread(key = area, value = n, fill = 0)  # Spread the 'area' column into multiple columns

# View the summary table
print(summary_table_tool_area)

@@
